---
title: "Draft_Report"
author: "Every Byte Helps"
date: "Wednesday, June 2, 2015"
output:
  html_document:
    toc: true
    number_sections: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
---

# Introduction

This document summarizes Every Byte Help's efforts to learn and analyze Colorado River Watch's water quality data.  [Colorado River Watch (RiverWatch)]( http://www.coloradoriverwatch.org/) is an organization that trains volunteer groups such as schools, envvironmental groups and fishing organization to correctly collect water quality samples.  These samples are analyzed using Colorado Department of Natural Resources facitlities. The data is stored in the [Colorado Data Sharing Network (CDSN)](http://www.coloradowaterdata.org/) database.

While River Watch has collected data at sites across the state, this initial task focuses on data collected in Roaring Fork Basin.  The Roaring Fork River is a tributary to the Upper Colorado River beginning above Aspen and joining the Colorado River at Glenwood Springs.  The Roaring Fork Basin was recommended due to the relatively high number of sampling locations with long period of records.


# Data

```{r dataRead, echo=FALSE, message=FALSE}
### load libraries
library(knitr)
opts_chunk$set(comment=FALSE, echo = FALSE,  message = FALSE)

library(kfigr, quietly = TRUE) 
library(ggplot2, quietly = TRUE)
library(dplyr, quietly = TRUE )
library(leaflet)


if (!requireNamespace('htmlwidgets') || packageVersion('htmlwidgets') <= '0.3.2')
  devtools::install_github('ramnathv/htmlwidgets')
# install DT
if (!require("DT")) devtools::install_github("rstudio/DT")
library(DT)


load("../data/binaryDat.Rdata")  ## created by data_processing.R ; binary for efficiency

load("../data/binaryFlow.Rdata") ## created by USGS Retreival

# A function for captioning and referencing images
fig <- local({
    i <- 0
    ref <- list()
    list(
        cap=function(refName, text) {
            i <<- i + 1
            ref[[refName]] <<- i
            paste("Figure ", i, ": ", text, sep="")
        },
        ref=function(refName) {
            ref[[refName]]
        })
})
```

Data used in this project was queried from the CDSN on April 30th.  The original query produced `r nrow(full_dat)` records from `r length(unique(full_dat$analyte))` analytes collected at `r length(unique(full_dat$Monitoring.Location.ID)) ` sites.  This data was collected between `r min(full_dat$date)` to `r max(full_dat$date)`.


## Samping Periods
```{r dataTable1, echo=FALSE }

datatable(events_by_sites[,-c(1,3,4) ],rownames = FALSE )

```


The following figures shows the dates of sampling events by site. From this plot, some observations are that routine sampling at many sites began around 2000 and sometime around 2004-5, sampling frequency changed at a number of locations.  River Watch indicated that this was done in part to conserve and reallocate resources.


```{r allEventPlot, fig.width=11, fig.height = 12, fig.cap=fig$cap("allEventPlot", "All Roaring Fork Sampling Events by date.")}

events_by_sites2 <- filter(events,date >= as.Date("1/1/1900", format = "%m/%d/%Y") )  %>% group_by(Monitoring.Location.ID )%>% 
    summarize( n_events = n_distinct(date),
               earliest = min(date), 
               latest = max(date)) %>% 
    arrange( desc(n_events))

### examine date range with plot

events %>% ggplot( aes(x = date, y = Monitoring.Location.Name )) + geom_point()

```


Based on `r fig$ref("allEventPlot")`, it was decided to initially focus on data collected a after January 1, 2000 and site with records collected during at least 10 years.  Events from this subset of sites are shown below.


```{r smallEventPlot, fig.width=11, fig.height = 6, fig.cap=fig$cap("smallEventPlot", "Dates of sampling events for reduced set of locations and events after January 1, 2000") }


### examine date range with plot

filter(events,date >= as.Date("1/1/1999", format = "%m/%d/%Y") & Monitoring.Location.ID %in% main_sites$Monitoring.Location.ID) %>% ggplot( aes(x = date, y = Monitoring.Location.Name )) + geom_point()

```

## Sampling Sites 

The following figure certainly suggests that there are  a few clusters of sites that might be evaluated to assess whether they could be considered as the same, thereby increasing representation along the main stem of the Roaring Fork and the Crystal River ( the southwestern tributary).



```{r siteFigure, fig.cap=fig$cap("siteFigure", "Sampling locations.  Blue pins indicate regularly monitored sites. Red points indicate infrequently or inconsistently monitored sites. ") }

good_sites <- unique( small_dat[ , names(small_dat) %in% c("Monitoring.Location.ID", "Monitoring.Location.Name", "Monitoring.Location.Latitude", "Monitoring.Location.Longitude"   ) ] ) 

all_sites <- unique( full_dat[ , names(small_dat) %in% c("Monitoring.Location.ID", "Monitoring.Location.Name", "Monitoring.Location.Latitude", "Monitoring.Location.Longitude"   ) ] ) 

names(good_sites)[3:4] <- c("latitude", "longitude")
names(all_sites)[3:4] <- c("latitude", "longitude")

m = leaflet() %>% addTiles()
m %>% addCircles(data=all_sites) %>% addMarkers(data=good_sites, popup=good_sites$Monitoring.Location.Name )


```

## Analytes

# Summary Statistics


## Tables
## Figures
## Maps

# Trends

To qualitatively explore trends, a linear regression was fit to the result for each site.  A shiny app was created allowing the user to select an analyte and to select whether the results should be log-transformed.  To allow a log transformation to be applied, all values reported as zero were set to the detection limit.  

*Currently this shiny app is hosted on shiny.io here.  Theoretically, we should be able to incorporate a shiny application into a markdown document.*

** [shinyApp](https://mjpdenver.shinyapps.io/RF_Trends) **

Scrolling through these plots, several items should be noted.

* For most metals, a high proportion of non-detects make linear models inappropriate.
* For some analytes, multiple detection limits were used.
* For some analytes such as sodium and potassium, it appear that earlier results were recorded as zero or below detection.  Given the subsequent higher levels, this is suspect and should be checked.
* For analytes like total copper, it appears that values are reported that are less than the dection limit.  It is not clear if this represents an inconsistency in the reporting procedures.

## Significant Trends
## Figures
## Maps

# Stream Segment Standards and Exceedances.
## Discussion of standards
## Table Value Standards

# Flow Data
## Location of USGS Gages

The following figure shows the locations of USGS Gages in the Roaring Fork Basin.  Markers indicates stations with nearly complete data records.  Red points indicate locations with intermediate flow records.  This may be due to the fact that these sites freeze up in the winter.  Check.


```{r usgslocation}

m = leaflet() %>% addTiles()
m %>% addCircles(data=usgs_site_info, col = "red" ) %>% 
     addMarkers(data = subset(usgs_site_info, major==TRUE), popup=paste(subset(usgs_site_info, major==TRUE)$station_nm, subset(usgs_site_info, major==TRUE)$site_no, sep = "-") )


```


## Correlation with Hardness



```{r cor_w_hardness}

sub_dat <- filter(small_dat, Characteristic.Name == "Hardness, Ca, Mg")
names(flow_wide)[7] <- "Glenwood"

left_join(sub_dat, flow_wide, by = "date",copy = TRUE )  %>% 
    ggplot(aes( x = Glenwood, y = Result.Value)) + labs( x = "Log(Flow) cfs", y = "log(Hardness) mg/L as CaO2") + 
    geom_point() + facet_wrap(~Monitoring.Location.Name, scales = "free_y") + 
    scale_x_log10() + scale_y_log10() + stat_smooth()

    ## adds flow columns #### 


```


# Next Steps
## Recommendations
## Other  basins

# Petty To-Do list

 1. Column width for DT datatable
2. Captions
