---
title: "RiverWatch"
output: html_document
---

The following information identifies:

1)  Median freq. of most active stations over the period 1990-11-26 through 2012-12-03
2)  The top 50% most active stations determine by number of samples over the noted period.  
3)  Histogram of each of station from the top 50% most frequented sites.,  to identify sampling consistency from year to year.
4)  The bottom 50% least active stations determine by number of samples over the noted period.
5)  Count by Org. Name over the period 1990-11-26 through 2012-12-03
6)  Each Organizations sampling counts by year.


NOTE:  Any records without data (NAs) were still included.

```{r}
library(plyr)
library(ggplot2)
# setwd("/Users/cyndikoehl/Documents/Kaggle and EBH/RiverWatch")
data <- read.csv("../data/processed/events.csv", colClasses = "character" )
stationData <- read.csv("../data/processed/stations.csv", colClasses = "character" )

data$calDate <- as.Date(data$date,format="%Y-%m-%d")
data$weekNo <- as.numeric(format(data$calDate,"%U"))
data$yearNo <- format(data$calDate,format="%Y")
data$monthNo <- format(data$calDate,format="%m")
data$yearMonthNo <- format(data$calDate,format="%Y-%m")


```

Median freq. of most active stations over the period 1990-11-26 through 2012-12-03


```{r}
countStn. <- count(data,c("Stn."))
countStn. <- countStn.[order(countStn.$freq,decreasing=TRUE),]
median(countStn.$freq)
```


The top 50% most active stations determine by number of samples over the noted period.  

```{r}
countStn.50sub <- subset(countStn.,countStn.$freq>= (median(countStn.$freq) ) )
data2 <- merge(x=countStn.50sub, y=stationData, by.x = "Stn.", by.y = "Station", all.x=TRUE )
countStn.50sub <- data2[,c(1:3)]
countStn.50sub <- countStn.50sub[order(countStn.50sub$freq,decreasing=TRUE),]
countStn.50sub
```

Histogram of each of station from the top 50% most frequented sites.,  to identify sampling consistency from year to year.

```{r}
for ( counter in 1:nrow(countStn.50sub)){

aStnByYear <-  subset(data,data$Stn. == countStn.50sub[counter,1] )
singleStationData <- subset(stationData,stationData$Station == countStn.50sub[counter,1] ) 
print(ggplot(aStnByYear, aes(x=c(yearNo))) +  
  geom_histogram(binwidth=0.5) + labs(x="Year",y="Number of Samples") + ggtitle(paste0("Station #",countStn.50sub[counter,1],": ",singleStationData$Name)) +
  theme(axis.text.x = element_text(color="#993333", size=8, angle=45)))

}

```

The bottom 50% least active stations determine by number of samples over the noted period.  

```{r}
countStn.50sub <- subset(countStn.,countStn.$freq< (median(countStn.$freq) ) )
data2 <- merge(x=countStn.50sub, y=stationData, by.x = "Stn.", by.y = "Station", all.x=TRUE )
countStn.50sub <- data2[,c(1:3)]
countStn.50sub <- countStn.50sub[order(countStn.50sub$freq,decreasing=TRUE),]
countStn.50sub
```

Count by Org. Name over the period 1990-11-26 through 2012-12-03

```{r}
countGroup <- count(data,c("Org.Name"))
countGroup <- countGroup[order(countGroup$freq,decreasing=TRUE),]
countGroup
```

Each Organizations sampling counts by year.

```{r}
for ( counter in 1:nrow(countGroup)){

aGroupByYear <-  subset(data,data$Org.Name == countGroup[counter,1] )
print(ggplot(aGroupByYear, aes(x=c(yearNo))) +  
  geom_histogram(binwidth=0.5) + labs(x="Year",y="Number of Samples") + ggtitle(paste0("Group Name ",countGroup[counter,1])) +
  theme(axis.text.x = element_text(color="#993333", size=8, angle=45)))

}

```